### 2017-01-11 robust-software.tex
Combined all references to documentation #38 . Moved non-documentation parts around the manuscript into appropriate sections.

\section{Have a README that briefly explains what the software does and what its dependencies are.}

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. 
Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards};
their key common features are listed below.

\textbf{Explain what the software does:} At the beginning of the README,
explain what the software does in one or two sentences.
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\textbf{List required dependencies:} Often, software depends on
specific versions of libraries, modules, or operating systems. 
This is
entirely reasonable as long as it is properly documented. Include the name and
version number for each dependency.  Several methods exist for acquiring
required packages: either provide the commands necessary to download the
dependencies; link to the software homepage; or (our preference) use a
dependency manager. We address dependencies in more detail in rule 5.

\textbf{Provide compilation/installation instructions:}
If the software needs to be compiled
or installed, list those instructions in the README. Also
mention if you recommend they use a pre-compiled binary instead
through a system such as pip, yum, or apt.

\textbf{List input and output files}: All possible input and output files
should be listed in this section. Do the files use a particular
standard, an extension of an existing format, or your own
format?  If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
(You get bonus points if you include a script to convert between
a standard format and your file format). If there is no rigorous format
(which is common with log files), show at least a few lines from an
example file and explain what the sections mean.

Input files and their formats are included in most documentation, but
intermediate, auxiliary, and log files are often missing.
\emph{All} files should be listed in the README, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. If your users might need to know,
``Does this program report the percentage of reads trimmed to remove
adapter sequences?'' they should be able to check the README and confidently
say, ``Yes, it is in the log file''.

\textbf{State attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work. These files should be available \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of ``expert'' users
that can be contacted if new users have problems with the software.

The README file for \withurl{khmer}{https://github.com/dib-lab/khmer/blob/master/README.rst}
is a good model:
it explains the software's purpose,
tells readers where to get help,
and includes links to a CITATION file (explaining how to cite the project)
and the license file.

\section{Tell the user what could be done and what was actually done.}

A robust program should
\textbf{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of the software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).  Long output is also much less likely to be read{\ldots}

Usage should provide all of the information necessary to run the
software. It is invoked either by running the software without
any arguments, running it with incorrect arguments, or by
explicitly choosing a help or usage option.
More standard command-line behaviours are detailed in \cite{Seemann2013}.

Almost all command-line applications use a combination of
POSIX~\cite{posix2016} and GNU~\cite{gnustandards} standards for usage. 
An example of good usage is GNU's \texttt{mkdir} command, which makes
new directories:

\begin{small}
\begin{verbatim}
$ mkdir --help
Usage: mkdir [OPTION]... DIRECTORY...
Create the DIRECTORY(ies), if they do not already exist.

Mandatory arguments to long options are mandatory for short options too.
  -m, --mode=MODE   set file mode (as in chmod), not a=rwx - umask
  -p, --parents     no error if existing, make parent directories as needed
  -v, --verbose     print a message for each created directory
  -Z, --context=CTX  set the SELinux security context of each created
                      directory to CTX
      --help     display this help and exit
      --version  output version information and exit

Report mkdir bugs to bug-coreutils@gnu.org
GNU coreutils home page: <http://www.gnu.org/software/coreutils/>
General help using GNU software: <http://www.gnu.org/gethelp/>
For complete documentation, run: info coreutils 'mkdir invocation'
\end{verbatim}
\end{small}

The key features of this usage message are:

@@

\item[\textbf{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more about the software.

@@

(And as a corollary
if the software is configurable in any way,
it should \textbf{check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two hours
that it isn't going to save its results
because the requested directory doesn't exist.)

@@

@@

\item[\textbf{Documentation}] We think it is very important for
  developers to document their work, but our experience is that people
  are unlikely do it during normal development, and we don't want to
  recommend practices that people won't actually adopt. However, it is worth
  noting that software that is widely used and contributed to has and enforces
  the need for good documentation~\cite{gentleman2004}.
  Numerous resources exist for writing high quality
  documentation~\cite{karimzadeh2016} and so we do not cover it here.

>>>>

\section{Document your code and usage}

Numerous resources exist for writing high quality
documentation~\cite{karimzadeh2016} and so we do not cover it here, except for
two minimal elements: the README and usage.

The README is the first stop for most new users, usually available even before
the software is installed.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. Usage is a terse, informative command-line help message that
guides the user in the correct use of the software.

Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards};
their key common features are listed below. 
The README file for \withurl{khmer}{https://github.com/dib-lab/khmer/blob/master/README.rst}
is a good model to emulate.

\item[\textbf{Explain what the software does.}] 
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

\item[\textbf{List required dependencies.}] Often, software depends on
specific versions of libraries, modules, or operating systems. Include the name and
version number for each dependency or a link to the requirements file. 
We address dependencies in more detail in rule 5.

\item[\textbf{Provide compilation/installation instructions:}]
If the software needs to be compiled
or installed, list those instructions in the README.

\item[\textbf{List input and output files.}] All possible input and output files
should be listed in this section. 
If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
If there is no rigorous format, as is common with log files, show at least a
few lines from an
example file and explain what the sections mean.

\emph{All} files should be listed, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. If your users might need to know,
``Does this program report the percentage of reads trimmed to remove
adapter sequences?'' they should be able to check the README and confidently
say, ``Yes, it is in the log file''.

\item[\textbf{State attributions and licensing.}] Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work.
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
\end{description}

In addition to a README, a robust program should
\textbf{print usage information when launching from the command line that explains the software's features}.
Usage information provides the first line of help for both first-time and
experienced users of command-line applications. Terseness is
important: usage that extends for multiple screens is 
difficult to read or refer to on the fly.

Usage is invoked either by running the software without
any arguments, running it with incorrect arguments, or by
explicitly choosing a help option.
More standard command-line behaviours are detailed in \cite{Seemann2013}.

Almost all command-line applications use a combination of
POSIX~\cite{posix2016} and GNU~\cite{gnustandards} standards for usage. 
The key features of a usage message are:

@@

\item[\textbf{Where to find more information}] Whether it's an email
  address, web site, or manual page, there should be an indication
  where the user can go to find out more.

@@

Documentation beyond the README and usage is up to the developer's discretion.
We think it is very important for developers to document their work, but our
experience is that people are unlikely do it during normal development, and we
don’t want to recommend practices that people won’t actually adopt. However, it
is worth noting that software that is widely used and contributed to has and
enforces the need for good documentation~\cite{gentleman2004}.

@@

\textbf{Check that all input values are in a reasonable range at startup}. 
Few things are as annoying as having a program announce after running for two
hours that it isn't going to save its results because the requested directory
doesn't exist.

@@
